---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
##Requirements
I use these packages. Install packages if you don't have these.
```{r results='hide'}
library(dplyr)
library(ggplot2)
library(grid)
library(RColorBrewer)
```
###Donwloading and loading data
In this step, I download the data from [here]("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip") and store it in data directory under current working directory(Check your working directory before download). And unzip it and load to data frame named activity.
```{r echo=TRUE}
if(!dir.exists("./data")){
    dir.create("./data")
}

fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if(!file.exists("./data/pa1zip.zip")){
    file.create("./data/pa1zip.zip")
}
download.file(fileUrl, "./data/pa1zip.zip")
unzip("./data/pa1zip.zip",exdir = "./data")
activity <- read.csv("./data/activity.csv")
```
The variables included in this dataset are:  
-**steps**: Number of steps taking in a 5-minute interval (missing values are coded as NA)  
-**date**: The date on which the measurement was taken in YYYY-MM-DD format  
-**interval**: Identifier for the 5-minute interval in which measurement was taken.
Let's see this dataset briefly(first 6 and last 6 rows).
```{r}
str(activity)
head(activity)
tail(activity)
```
As you can see, there are NAs in variable *steps*. And variable *date* is loaded as factor not date. Now let's check how many NAs there are in *steps*.

```{r}
sum<-sum(is.na(activity$steps))
sum
sum/length(activity$steps)
```
There are `r sum` NAs in the *steps* and it is about `r 100*sum/length(activity$steps)`% of rows. For further analysis I make data.frame **complete_activity** by removing NAs of *steps*.
```{r}
complete_activity <- activity[complete.cases(activity$steps),]
```

## What is mean total number of steps taken per day?
I make histogram of the total number of steps taken each day.  
```{r}
total_steps <- select(activity,steps,date) %>% group_by(date)%>% summarise(steps = sum(steps,na.rm = TRUE))
mean_steps <- mean(total_steps$steps)
median_steps <- median(total_steps$steps)
cols <- brewer.pal(3,'Dark2')
mean_string<-paste("mean =",round(mean_steps,2))
median_string<-paste("median =",round(median_steps,2))

g <- ggplot(total_steps,aes(steps))
g + geom_histogram(binwidth = 800) + 
    labs(title = 'The Total Number of Steps\nTaken Each Day', 
         x = 'The total number of steps') + 
    geom_vline(xintercept = mean_steps,lwd = 2,col = cols[1]) +
    geom_vline(xintercept = median_steps,lwd = 2,col = cols[2]) +
    annotate("text",x=5000,y=9.0,label = mean_string,col = cols[1],size = 4) +
    annotate("text",x=5000,y=7.5,label = median_string,col = cols[2],size = 4)+
    theme(plot.title = element_text(size = 28,lineheight=.8,vjust=3))
```

In this histogram, the mean total number of steps taken per day is `r mean_steps`. And a number of days that the total number of steps is less than 800 is 10. It is same as a number of days of walking 10000-10800 steps in a day.


## What is the average daily activity pattern?
I make a time series plot of the 5-minute interal and the average number of steps taken, averaged across all days.
```{r}
steps_pattern <- group_by(complete_activity,interval)%>%summarise(steps = mean(steps))
max_interval<-steps_pattern[steps_pattern$steps==max(steps_pattern$steps),1]
maxg1 <- ggplot(steps_pattern,aes(interval,steps))
g1 + geom_line(aes(interval,steps)) + labs(title = 'The average number of steps of the 5-minute interval',
                                           subtitle = 'averaged across all days', x = '5-minute interval',
                                           y = 'Average number of steps') +
    geom_point(aes(max_interval,max(steps_pattern$steps)),lwd=3,col = cols[3])+
    annotate("text",x=500,y=max(steps_pattern$steps),
             label = paste("maximum=",round(max(steps_pattern$steps),3)))
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
